Implementing Capon Adaptive Beamforming on the GPU for Real Time Cardiac Ultrasound Imaging

### 2012-06-05 HRUIonGPU/jrnl_article.tex
tuffc capon on speed

\subsection{Calculating Multiple Sample Covariance Matrices}
The straight forward sum in (\ref{eq:R}) has complexity $O(L^2KN) = O(L^3)$ if $L$ is a small fraction of $M$. This is actually the same as for inverting a $L \times L$ matrix. However, little attention has been devoted to this step in the literature until now. Note that for a typical array configuration, N is usually small compared to L and M. Hence, for smaller $L$ the construction of $\mat{\hat{R}}$ is dominated by $M$, since $K = M - L + 1$. From (\ref{eq:R}) it is obvious that calculation of one element in $\mat{\hat{R}}$ is independent of all the others. A natural level of granularity would then be to assign one thread to each element of $\mat{\hat{R}}$. This will minimize global reads and writes needed per thread, but unfortunately the maximum number of threads per block puts an upper limit on $L$, $L_{max}=32$, if we want to avoid block-to-block synchronization.

@@

For the results presented in this paper, we have used an unreleased GPU implementation of Gauss Jordan (GJ) elimination (by NVIDIA) to solve $\mat{R}\vec{b} = \vec{1}$, where $\vec{b} = \mat{R}^{-1}\vec{1}$, but we will go through an implementation of an $\mat{U}^H\mat{D}\mat{U}$-based solver. ...

@@

@@

\subsection{Compute Beamformer Output}
One thread per weight ($L$ threads). Let one thread sum the solution vector $\vec{b}$, or let several threads cooperate to form the sum using a binary reduction sum.
Let one thread read one element of $\vec{b}$, scale it with the inverse sum, and optionally write it back to global memory. The final output can be formed by either reduce the data down to length $L$, or increase the weight vector to length $M$. Since we have decided to have $L$ threads per output, the data is reduced to length $L$ by calculating $x_{r} = \sum_{k=r}^{r+K-1}x_{k}$ for $r \in [0, L-1]$. Each thread now computes $z_r = w_r^*x_r$, and finally one thread computes and outputs $z = \sum_{r=0}^{L-1}$ to global memory.

@@

\subsection{Iterative Capon}
The idea behind iterative Capon is to perform Capon using small subarrays in several iterations until there is only two subarrays to sum. In this way we fix the size of L, hence the complexity of solving $\mat{R}\vec{b} = \vec{1}$. However, for each iteration we re-obtain L-1 degrees of freedom until all are retrieved after $K = M-L+1$ iterations. Since subarray averaging has been applied to decorrelated the signal, we might loose this effect when reestablishing all degrees of freedom. The number of iterations should therefore be restricted to $K/2$ to equal our standard choice of $L = M/2$ when it comes to degrees of freedom.

@@

>>>>

\subsection{Calculating Multiple Sample Covariance Matrices}
The straight forward sum in (\ref{eq:R}) has complexity $O(L^2KN) = O(L^3)$ if $L$ is a small fraction of $M$. This is actually the same as for inverting a $L \times L$ matrix. However, little attention has been devoted to this step in the literature until now. Note that for a typical array configuration, N is usually small compared to L and M. Hence, for a small $L$ the construction of $\mat{\hat{R}}$ is dominated by $M$, since $K = M - L + 1$. From (\ref{eq:R}) it is obvious that calculation of one element in $\mat{\hat{R}}$ is independent of all the others. A natural level of granularity would then be to assign one thread to each element of $\mat{\hat{R}}$. This will minimize global reads and writes needed per thread, but unfortunately the maximum number of threads per block puts an upper limit on $L$, $L_{max}=32$, if we want to avoid block-to-block synchronization.

@@

For the results presented in this paper, we have used an unreleased GPU implementation of Gauss Jordan (GJ) elimination (by NVIDIA) to solve $\mat{R}\vec{b} = \vec{1}$, where $\vec{b} = \mat{R}^{-1}\vec{1}$. However we will go through an implementation of an $\mat{U}^H\mat{D}\mat{U}$-decomposition-based solver to point out challenges when implementing batched solving of small linear systems on the GPU. ... (Unsure if this should be included).

@@

For the smallest systems a direct solver using one thread per system is preferable. For this purpose we have implemented Cramer's rule for $3 \times 3$ (, $2 \times 2$?  and $4 \times 4$?) matrices.

@@

\subsection{Compute Beamformer Output}
(Still just notes) One thread per weight ($L$ threads). Let one thread sum the solution vector $\vec{b}$, or let several threads cooperate to form the sum using a binary reduction sum.
Let one thread read one element of $\vec{b}$, scale it with the inverse sum, and optionally write it back to global memory. The final output can be formed by either reduce the data down to length $L$, or increase the weight vector to length $M$. Since we have decided to have $L$ threads per output, the data is reduced to length $L$ by calculating $x_{r} = \sum_{k=r}^{r+K-1}x_{k}$ for $r \in [0, L-1]$. Each thread now computes $z_r = w_r^*x_r$, and finally one thread computes and outputs $z = \sum_{r=0}^{L-1}$ to global memory. The last sum can also be performed using a binary reduction sum across several threads for increased speed.

@@

\subsection{Iterative Capon}
The idea behind iterative Capon is to perform Capon using small subarrays in several iterations until there is only two subarrays to sum. In this way we fix the size of L, hence the complexity of solving $\mat{R}\vec{b} = \vec{1}$. However, for each iteration we re-obtain L-1 degrees of freedom until all are retrieved after $K = M-L+1$ iterations. Since subarray averaging has been applied to decorrelated the signal, we might loose this effect when reestablishing all degrees of freedom. The number of iterations should therefore be restricted to $K/2$ to equal the normal choice of $L = M/2$ when it comes to degrees of freedom.

@@

Describe the base band (IQ-demodulation) process that the data has been undergoing.
